### 2023-06-15 state-resilience.tex
up

\problemx{Downward-reachability problem}
{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and a downward-closed set $D
\subseteq S$.}
% {$\exists s' \in D ~ s \to^* s'$? \newline}\alain{ecrire $s  \to^* D$}
{$s  \to^* D$? \newline}

%\"Ozkan
\begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
For finite-branching WSTS%with strong compatibility
, a basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
%i 			.e. given a state $s$ of a WSTS
% in the regarded class 
%with strong compatibility 
%and a decidable downward-closed set $D$, it can be decided whether $\exists s' \in D ~ s \to^* s'$. 
\end{proposition}

% \alain{rappeler l'idée de la preuve}
The idea behind the proof is the following. For deciding whether a downward-closed set $D$ is reachable from $s$, one checks whether
$B_{\uparrow \post^*(s)} \cap D = \emptyset$, 
where $B_{\uparrow \post^*(s)}$ is a basis of $\uparrow \post^*(s)$,
%\alain{rapeler $B_{\uparrow \post^*(s)}$} 
 that is equivalent to $\post^*(s)\cap D = \emptyset$ by
Lemma~\ref{Lemma intersection}. For the converse direction, one computes the sequence of upward-closed sets
$U_n = \uparrow \post^{\leq n}(s)$ until it becomes stationnary. 
Decidability of downward-reachability leads to the decidability of the following stop condition:
asking whether $S \setminus U_n$ is reachable from $s$.

% \mathieu{Quelque chose de plus à dire sur la charactérisation?}

More concretely,
VASS for instance are $\uparrow \post^*$-effective WSTS \cite{DBLP:journals/corr/abs-2108-00889}. 
It is well-known that 
% Petri nets are WSTS with strong  compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}). 
% Since 
VASS are WSTS with strong compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that {\sc Bounded-state-resilience} is decidable for VASS.
Hence {\sc State-resilience} is decidable for %both Petri nets and 
VASS.
However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework. Moreover, we may show:

\begin{proposition}
There exist classes of WSTSs with strong 
%(upward)
 compatibility for which there don't exist an algorithm computing a basis of $\uparrow \post^*$.
\end{proposition}

@@

Results synthesis in the case $\Safe = \uparrow \Safe$:
\begin{center}
\begin{tabular}{ | l | c | c | c | c |}
\hline  Hypothesis & strong %upward 
			compatibility ~ & $\uparrow \post^*$ effective 
			& strong %upward 
					compatibility \textit{+} $\uparrow \post^*$ effective \\ \hline
   SRP & Undecidable (Thm~\ref{srp up down}) & Decidable (Thm~\ref{post srp})  & Decidable (Thm~\ref{postcomputable})\\ \hline
   BSRP & Undecidable (Thm~\ref{srp up down}) &  ??  & Decidable (Thm~\ref{ref ozkan}) \\ \hline
      kSRP & Undecidable (Thm~\ref{srp up down}) & ?? & Decidable (Thm~\ref{ref ozkan}) \\ \hline
 \end{tabular}
\end{center}

>>>>

\problemx{Downward-reachability problem}
{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and a downward-closed set $D
\subseteq S$.}
{$s  \to^* D$? \newline}

%\"Ozkan
\begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
For finite-branching WSTS%with strong compatibility
, a basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
\end{proposition}

The idea behind the proof is the following. For deciding whether a downward-closed set $D$ is reachable from $s$, one checks whether
$B_{\uparrow \post^*(s)} \cap D = \emptyset$, 
where $B_{\uparrow \post^*(s)}$ is a basis of $\uparrow \post^*(s)$,
%\alain{rapeler $B_{\uparrow \post^*(s)}$} 
 that is equivalent to $\post^*(s)\cap D = \emptyset$ by
Lemma~\ref{Lemma intersection}. For the converse direction, one computes the sequence of upward-closed sets
$U_n = \uparrow \post^{\leq n}(s)$ until it becomes stationnary. 
Decidability of downward-reachability leads to the decidability of the following stop condition:
asking whether $S \setminus U_n$ is reachable from $s$.

More concretely,
VASS for instance are $\uparrow \post^*$-effective WSTS \cite{DBLP:journals/corr/abs-2108-00889}. 
It is well-known that 
VASS are WSTS with strong compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that {\sc Bounded-state-resilience} is decidable for VASS.
Hence {\sc State-resilience} is decidable for %both Petri nets and 
VASS.
However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework. Moreover, we may show:

\begin{proposition}
There exist classes of WSTSs with strong 
 compatibility for which there doesn't exist an algorithm computing a basis of $\uparrow \post^*$.
\end{proposition}

@@

{\bf Results synthesis in the case $\Safe = \uparrow \Safe$}
\begin{center}
\begin{tabular}{ | l | c | c | c | c |}
\hline  Hypothesis & strong %upward 
			compatibility ~ & $\uparrow \post^*$ effective 
			& strong %upward 
					compatibility \textit{+} $\uparrow \post^*$ effective \\ \hline
   SRP & Undecidable (Thm~\ref{srp up down}) & Decidable (Thm~\ref{post srp})  & Decidable (Thm~\ref{postcomputable})\\ \hline
   BSRP & Undecidable (Thm~\ref{srp up down}) &  ??  & Decidable (Thm~\ref{ref ozkan}) \\ \hline
      kSRP & Undecidable (Thm~\ref{srp up down}) & ?? & Decidable (Thm~\ref{ref ozkan}) \\ \hline
 \end{tabular}
\end{center}
